---
title: "proportion_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Loading required packages
library(tidyverse) # for a variety of things including %>%

library(pheatmap) # for heatmap
library(heatmaply) # for heatmap
library(ggpubr) # for plotting

library(Rtsne) # for t-SNE

library(caret) # for fitting models
# library(class) # for knn
library(e1071) # for svm
library(randomForest) # for randomforest
library(janitor) # for cleaning names for randomforest
library(xgboost) # for xgboost

library(Hotelling) # for testing
```


# Preparing Data

## Reading Data

Note that the data is extremely large, so reading the data takes a long time. 

```{r reading data, eval = FALSE}

# batch1_2 = data.table::fread("NewData/batch3_4.csv")
# batch3_4 = data.table::fread("NewData/batch3_4.csv")
# batch5 = data.table::fread("NewData/batch5.csv")
# batch6 = data.table::fread("NewData/batch6.csv")
# batch7 = data.table::fread("NewData/batch7.csv")
# batch8_9 = data.table::fread("NewData/batch8_9.csv")
```

## Cleaning Data

Using a separate R script function.

```{r cleaning data, eval = FALSE}

# ## Loading function
# source("functions/proportion_analysis_data_clean_fn.R")
# 
# ## Cleaning data
# b12 = data_clean(batch1_2)
# b34 = data_clean(batch3_4)
# b5 = data_clean(batch5)
# b6 = data_clean(batch6)
# b7 = data_clean(batch7)
# b89 = data_clean(batch8_9)
```

## Saving New Data

```{r saving data, eval = FALSE}

# ## Saving as RDS files
# saveRDS(b12, file = "rdata/batch1_2.rds")
# saveRDS(b34, file = "rdata/batch3_4.rds")
# saveRDS(b5, file = "rdata/batch5.rds")
# saveRDS(b6, file = "rdata/batch6.rds")
# saveRDS(b7, file = "rdata/batch7.rds")
# saveRDS(b89, file = "rdata/batch8_9.rds")
```

## Loading Data

```{r}

## Loading data
b12 = readRDS("rdata/batch1_2.rds")
b34 = readRDS("rdata/batch3_4.rds")
b5 = readRDS("rdata/batch5.rds")
b6 = readRDS("rdata/batch6.rds")
b7 = readRDS("rdata/batch7.rds")
b89 = readRDS("rdata/batch8_9.rds")

```


## Obtaining Barcodes

These barcodes indicate what cell type markers are expressed, for example:

- 00000000 = no cell markers at all
- 00001000 = cd16+ non-macrophages (cd68-) 
- 00000101 = all pd-l1/pd-l1 non-melanoma  --> if expressing multiple cell type markers, these types are separated by a "/"

```{r obtaining barcodes}
# 
# celltype_cols = c("melanoma (sox10+)",
#                   "cytotoxic t-cell (cd8+)",
#                   "macrophage (cd68+)",
#                   "cd16+ macrophages (cd68+)",
#                   "cd16+ non-macrophages (cd68-)",
#                   "all pd-l1",
#                   "pd-l1 melanoma",
#                   "pd-l1 non-melanoma")
# 
# ## Sourcing in get_barcode function
# source("functions/get_barcode_fn.R")
# 
# 
# ## Obtaining barcoded datasets
# b12_barcode = b12 %>% 
#   mutate(celltype = get_barcode(b12))
# b12_barcode = b12_barcode %>% select(-celltype_cols)
# 
# b34_barcode = b34 %>% 
#   mutate(celltype = get_barcode(b34))
# b34_barcode = b34_barcode %>% select(-celltype_cols)
# 
# b5_barcode = b5 %>% 
#   mutate(celltype = get_barcode(b5))
# b5_barcode = b5_barcode %>% select(-celltype_cols)
# 
# b6_barcode = b6 %>% 
#   mutate(celltype = get_barcode(b6))
# b6_barcode = b6_barcode %>% select(-celltype_cols)
# 
# b7_barcode = b7 %>% 
#   mutate(celltype = get_barcode(b7))
# b7_barcode = b7_barcode %>% select(-celltype_cols)
# 
# b89_barcode = b89 %>% 
#   mutate(celltype = get_barcode(b89))
# b89_barcode = b89_barcode %>% select(-celltype_cols)
#
#
## Saving as RDS files
# saveRDS(b12_barcode, file = "rdata/b12_barcode.rds")
# saveRDS(b34_barcode, file = "rdata/b34_barcode.rds")
# saveRDS(b5_barcode, file = "rdata/b5_barcode.rds")
# saveRDS(b6_barcode, file = "rdata/b6_barcode.rds")
# saveRDS(b7_barcode, file = "rdata/b7_barcode.rds")
# saveRDS(b89_barcode, file = "rdata/b89_barcode.rds")
```

## Obtaining Proportions

```{r getting proportions of barcodes, eval = FALSE}

# ## Sourcing in get_barcode_prop function
# source("functions/get_barcode_prop_fn.R")
# 
# ## Obtaining barcode proportions
# b12_barcode = readRDS("rdata/b12_barcode.rds")
# b12_barcodeprop = get_barcode_prop(b12_barcode)
# 
# b34_barcode = readRDS("rdata/b34_barcode.rds")
# b34_barcodeprop = get_barcode_prop(b34_barcode)
# 
# b5_barcode = readRDS("rdata/b5_barcode.rds")
# b5_barcodeprop = get_barcode_prop(b5_barcode)
# 
# b6_barcode = readRDS("rdata/b6_barcode.rds")
# b6_barcodeprop = get_barcode_prop(b6_barcode)
# 
# b7_barcode = readRDS("rdata/b7_barcode.rds")
# b7_barcodeprop = get_barcode_prop(b7_barcode)
# 
# b89_barcode = readRDS("rdata/b89_barcode.rds")
# b89_barcodeprop = get_barcode_prop(b89_barcode)
# 
# ## Combining batches
# barcodeprop_dat = bind_rows(b12_barcodeprop,
#                             b34_barcodeprop,
#                             b5_barcodeprop,
#                             b6_barcodeprop,
#                             b7_barcodeprop,
#                             b89_barcodeprop)
# 
# ## Obtaining and cleaning clinical data
# clinical = readxl::read_xlsx("NewData/clinical.xlsx")
# 
# clinical_toexclude = clinical$`Comments/May want to exclude from analysis` %>% 
#   replace_na("none") %>% 
#   str_detect("Little melanoma|DAPI") # Removing all patients that have little melanoma or failed DAPI stain
# 
# clinical = clinical[!clinical_toexclude, ] %>% 
#   select("imageid" = "Pathology Reference No",
#          "response" = "Response",
#          "treatment" = "Treatment")
# 
# ## Adding response from clinical data
# barcodeprop_dat = clinical %>%
#   inner_join(barcodeprop_dat, by = "imageid") # inner_join to contain all of the patients with both clinical and cellprop data
# 
# 
# ## Recoding barcodes to something more meaningful
# 
# ## Sourcing in function to get barcodes
# source("functions/get_barcodename_fn.R")
# 
# ## Obtaining barcode names
# barcodename_vec = sapply(unique(barcodeprop_dat$celltype), get_barcodename) %>%
#   setNames(unique(barcodeprop_dat$celltype), .)
# 
# levels(barcodeprop_dat$celltype) = names(barcodename_vec[match(levels(barcodeprop_dat$celltype), barcodename_vec)]) # Matching the current levels with barcodename_vec
# # and obtaining corresponding description
# 
# 
# # Changing to wide format
# barcodeprop_dat = barcodeprop_dat %>%
#   pivot_wider(id_cols = c(1, 2, 3, 7),
#               names_from = "celltype",
#               values_from = "prop")
# 
# ## Duplicated patient in batch 1&2 and batch 7 --> Tuba said to use batch 7
# barcodeprop_dat = barcodeprop_dat %>%
#   filter(imageid != "SP-15-016503" | (imageid == "SP-15-016503" & batch == "b7"))
# 
# ## Saving
# saveRDS(barcodeprop_dat, file = "rdata/barcodeprop_dat.rds")
```

```{r loading barcodeprop data}

## Loading barcodeprop data
barcodeprop_dat = readRDS("rdata/barcodeprop_dat.rds")

## Coercing response into a factor variable
barcodeprop_dat$response = as.factor(barcodeprop_dat$response)
```


# Proportions Analysis

Note that the patients are split into those with combination anti-PD-1 + anti-CTLA-4 therapy; and those with anti-PD-1 monotherapy. We will be performing subsequent analysis on each group individually.

```{r splitting by treatment}

# ## Combination
# barcodeprop_comb = barcodeprop_dat %>% 
#   filter(treatment == "combination anti-PD-1 + anti-CTLA-4") %>% 
#   select(-treatment)
# 
# ## Monotherapy
# barcodeprop_mono = barcodeprop_dat %>% 
#   filter(treatment == "anti-PD-1 monotherapy") %>% 
#   select(-treatment)
# 
# ## Changing response to factor
# barcodeprop_comb$response = as.factor(barcodeprop_comb$response)
# barcodeprop_mono$response = as.factor(barcodeprop_mono$response)
# 
# ## Saving
# saveRDS(barcodeprop_comb, file = "rdata/barcodeprop_comb.rds")
# saveRDS(barcodeprop_mono, file = "rdata/barcodeprop_mono.rds")
# 
## Loading barcodeprop data
barcodeprop_comb = readRDS("rdata/barcodeprop_comb.rds")
barcodeprop_mono = readRDS("rdata/barcodeprop_mono.rds")
```

## Proportion Heatmap {.tabset}

### Combination therapy

```{r combination heatmap, fig.height = 10, fig.width = 10}

# Matrix


comb_barcodeprop_mat =
  barcodeprop_comb %>% 
  select(-imageid, -response, -batch) %>% 
  as.matrix() %>% 
  t() %>% 
  
  # ## Sqrt transformation
  # apply(2, sqrt)
  
  ## Changing column names
  `colnames<-`(barcodeprop_comb$imageid)



# Annotations


## Column annotation: response and batch

comb_anncol = data.frame("Response" = barcodeprop_comb$response,
                         "Batch" = barcodeprop_comb$batch) %>% 
  `rownames<-`(colnames(comb_barcodeprop_mat)) # Setting rownames to imageid (colnames of barcodeprop_mat)


## Row annotation: melanoma

comb_annrow = data.frame("Cell Type" = rep("Non-melanoma", nrow(comb_barcodeprop_mat))) %>%  # Populating with all Non-melanoma for now
  `colnames<-`("Cell Type") %>%  # For nicer plotting
  `rownames<-`(rownames(comb_barcodeprop_mat))  # Setting rownames to cell types (rownames of cellprop_mat)

## Obtaining row indexes which are "melanoma" AND NOT "non-melanoma"
melanoma_index = (str_detect(rownames(comb_annrow), "melanoma") & !str_detect(rownames(comb_annrow), "non-melanoma"))

comb_annrow[melanoma_index, ] = "Melanoma" # Changing melanoma cells to melanoma classification



# Creating heatmap


comb_heat = ggheatmap(comb_barcodeprop_mat,
                      scale = "row",
                      dist_method = "euclidean",
                      hclust_method = "complete",
                      dendrogram = "column",
                      main = "Combination Therapy Cell Proportions",
                      col_side_colors = comb_anncol,
                      row_side_colors = comb_annrow,
                      showticklabels = c(FALSE, TRUE))

comb_heat_interactive = heatmaply(comb_barcodeprop_mat,
                                  scale = "row",
                                  dist_method = "euclidean",
                                  hclust_method = "complete",
                                  dendrogram = "column",
                                  main = "Combination Therapy Cell Proportions",
                                  col_side_colors = comb_anncol,
                                  row_side_colors = comb_annrow,
                                  showticklabels = c(FALSE, TRUE))




comb_heat
comb_heat_interactive

# ggsave("figures/combination_heatmap.jpg", plot = comb_heat, height = 10, width = 20)
```

### Monotherapy

```{r monotherapy heatmap, fig.height = 10, fig.width = 10}

# Matrix


mono_barcodeprop_mat =
  barcodeprop_mono %>% 
  select(-imageid, -response, -batch) %>% 
  as.matrix() %>% 
  t() %>% 
  
  # ## Sqrt transformation
  # apply(2, sqrt)
  
  ## Changing column names
  `colnames<-`(barcodeprop_mono$imageid)



# Annotations


## Column annotation: response and batch

mono_anncol = data.frame("Response" = barcodeprop_mono$response,
                         "Batch" = barcodeprop_mono$batch) %>% 
  `rownames<-`(colnames(mono_barcodeprop_mat)) # Setting rownames to imageid (colnames of barcodeprop_mat)


## Row annotation: melanoma

mono_annrow = data.frame("Cell Type" = rep("Non-melanoma", nrow(mono_barcodeprop_mat))) %>%  # Populating with all Non-melanoma for now
  `colnames<-`("Cell Type") %>%  # For nicer plotting
  `rownames<-`(rownames(mono_barcodeprop_mat))  # Setting rownames to cell types (rownames of cellprop_mat)

## Obtaining row indexes which are "melanoma" AND NOT "non-melanoma"
melanoma_index = (str_detect(rownames(mono_annrow), "melanoma") & !str_detect(rownames(mono_annrow), "non-melanoma"))

mono_annrow[melanoma_index, ] = "Melanoma" # Changing melanoma cells to melanoma classification



# Creating heatmap


mono_heat = ggheatmap(mono_barcodeprop_mat,
                      scale = "row",
                      dist_method = "euclidean",
                      hclust_method = "complete",
                      dendrogram = "column",
                      main = "Monotherapy Cell Proportions",
                      col_side_colors = mono_anncol,
                      row_side_colors = mono_annrow,
                      showticklabels = c(FALSE, TRUE))

mono_heat_interactive = heatmaply(mono_barcodeprop_mat,
                                  scale = "row",
                                  dist_method = "euclidean",
                                  hclust_method = "complete",
                                  dendrogram = "column",
                                  main = "Monotherapy Cell Proportions",
                                  col_side_colors = mono_anncol,
                                  row_side_colors = mono_annrow,
                                  showticklabels = c(FALSE, TRUE))




mono_heat
mono_heat_interactive

# ggsave("figures/monotherapy_heatmap.jpg", plot = mono_heat, height = 10, width = 20)
```

## Proportion PCA {.tabset}

### Combination therapy

```{r combination pca}

## PCA
comb_pca = prcomp(t(comb_barcodeprop_mat), scale = TRUE)

## Plot data
comb_pca_df = tibble(
  PC1 = comb_pca$x[, 1],
  PC2 = comb_pca$x[, 2],
  PC3 = comb_pca$x[, 3],
  Response = barcodeprop_comb$response
)

## Plotting PCA
comb_pca1 = comb_pca_df %>% 
  ggplot() +
  aes(x = PC1, y = PC2, colour = Response) +
  geom_point() +
  labs(title = "Combination PC1 ~ PC2") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

comb_pca2 = comb_pca_df %>% 
  ggplot() +
  aes(x = PC1, y = PC3, colour = Response) +
  geom_point() +
  labs(title = "Combination PC1 ~ PC3") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

comb_pca3 = comb_pca_df %>% 
  ggplot() +
  aes(x = PC2, y = PC3, colour = Response) +
  geom_point() +
  labs(title = "Combination PC2 ~ PC3") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

ggpubr::ggarrange(comb_pca1, comb_pca2, comb_pca3,
                  ncol = 3, common.legend = TRUE, legend = "bottom")
```

There isn't much separation in the PCA plots.

### Monotherapy

```{r monotherapy pca}

## PCA
mono_pca = prcomp(t(mono_barcodeprop_mat), scale = TRUE)

## Plot data
mono_pca_df = tibble(
  PC1 = mono_pca$x[, 1],
  PC2 = mono_pca$x[, 2],
  PC3 = mono_pca$x[, 3],
  Response = barcodeprop_mono$response
)

## Plotting PCA
mono_pca1 = mono_pca_df %>% 
  ggplot() +
  aes(x = PC1, y = PC2, colour = Response) +
  geom_point() +
  labs(title = "Monotherapy PC1 ~ PC2") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

mono_pca2 = mono_pca_df %>% 
  ggplot() +
  aes(x = PC1, y = PC3, colour = Response) +
  geom_point() +
  labs(title = "Monotherapy PC1 ~ PC3") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

mono_pca3 = mono_pca_df %>% 
  ggplot() +
  aes(x = PC2, y = PC3, colour = Response) +
  geom_point() +
  labs(title = "Monotherapy PC2 ~ PC3") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

ggpubr::ggarrange(mono_pca1, mono_pca2, mono_pca3,
                  ncol = 3, common.legend = TRUE, legend = "bottom")
```

There isn't much separation in the PCA plots.

## Cell type proportions $t$-SNE {.tabset}

### Combination therapy

```{r combination tsne, fig.height = 10, fig.width = 10}

## Initialising list to store plots
comb_tsne_list = list()

## Setting random seed for reproducibility
set.seed(2021)

## For loop for different perplexity values
for(perplexity in c(2, 5, 10)){
  
  ## Performing t-SNE
  comb_tsne = t(comb_barcodeprop_mat) %>% 
    Rtsne(perplexity = perplexity)
  
  ## Creating data for plotting
  comb_tsne_plotdata = tibble(
    `Dimension 1` = comb_tsne$Y[, 1],
    `Dimension 2` = comb_tsne$Y[, 2],
    `Response` = barcodeprop_comb$response
  )
  
  ## Plotting
  comb_tsne_list[[match(perplexity, c(2, 5, 10))]] = comb_tsne_plotdata %>% 
    ggplot() +
    aes(x = `Dimension 1`, y = `Dimension 2`, colour = `Response`) +
    geom_point() +
    ggtitle(paste("Combination Therapy t-SNE: Perplexity =", perplexity)) +
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5))
  
}

gridExtra::grid.arrange(comb_tsne_list[[1]],
                        comb_tsne_list[[2]],
                        comb_tsne_list[[3]],
                        ncol = 2)
```

There isn't much separation in the $t$-SNE plots either.

### Monotherapy

```{r monotherapy tsne, fig.height = 10, fig.width = 10}

## Initialising list to store plots
mono_tsne_list = list()

## Setting random seed for reproducibility
set.seed(2021)

## For loop for different perplexity values
for(perplexity in c(2, 5, 10)){
  
  ## Performing t-SNE
  mono_tsne = t(mono_barcodeprop_mat) %>% 
    Rtsne(perplexity = perplexity)
  
  ## Creating data for plotting
  mono_tsne_plotdata = tibble(
    `Dimension 1` = mono_tsne$Y[, 1],
    `Dimension 2` = mono_tsne$Y[, 2],
    `Response` = barcodeprop_mono$response
  )
  
  ## Plotting
  mono_tsne_list[[match(perplexity, c(2, 5, 10))]] = mono_tsne_plotdata %>% 
    ggplot() +
    aes(x = `Dimension 1`, y = `Dimension 2`, colour = `Response`) +
    geom_point() +
    ggtitle(paste("Monotherapy t-SNE: Perplexity =", perplexity)) +
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5))
  
}

gridExtra::grid.arrange(mono_tsne_list[[1]],
                        mono_tsne_list[[2]],
                        mono_tsne_list[[3]],
                        ncol = 2)
```

There isn't much separation in the $t$-SNE plots either.

## Classification {.tabset}

### Combination therapy

```{r combination data balance}

table(barcodeprop_comb$response)
```

The data is quite imbalanced.


```{r combination training accuracy}

## Initialising lists to store balanced accuracies
glm_acc = glm_balacc = rf_acc = rf_balacc = svm_acc = svm_balacc = xgb_acc = xgb_balacc = list()

set.seed(2021)

for(repeats in 1:20){
  
  ## Defining folds to use
  cvIndex = caret::createFolds(barcodeprop_comb$response, k = 10, returnTrain = TRUE)
  
  ## Initialising list of confusion matrices
  glm_conmat = rf_conmat = svm_conmat = xgb_conmat = list()
  
  for(fold in 1:5){
    
    ## Training data
    train_dat = barcodeprop_comb %>% .[cvIndex[[fold]], ] %>% select(-imageid, -batch)
    train_x = train_dat %>% select(-response)
    train_y = train_dat %>% pull(response)
    
    ## Testing data
    test_dat = barcodeprop_comb %>% .[-cvIndex[[fold]], ] %>% select(-imageid, -batch)
    test_x = test_dat %>% select(-response)
    test_y = test_dat %>% pull(response) 
    
    
    
    ## Fitting models
    glm_model = glm(response ~ ., 
                    data = train_dat, 
                    family = binomial(link = "logit"))
    
    rf_model = randomForest::randomForest(response ~ .,
                                          data = janitor::clean_names(train_dat)) # Need to clean names for randomforest to work
    
    svm_model = svm(train_x, train_y)
    
    xgb_model = xgboost(data = unlist(train_x) %>% matrix(nrow = nrow(train_x)),
                        label = as.numeric(train_y) - 1, # xgb needs label to be numeric 0 < y < 1
                        max_depth = 2, eta = 1, nthread = 1, nrounds = 2, objective = "binary:logistic")
    
    
    ## Predictions on test data
    glm_pred = predict(glm_model, test_x, type = "response") %>% 
      {ifelse(. > 0.5, "Responder", "Non-responder")} %>%  # Cutoff at probability =  0.5
      as.factor()
    
    rf_pred = predict(rf_model, janitor::clean_names(test_x), type = "response")
    
    svm_pred = predict(svm_model, test_x)
    
    xgb_pred = predict(xgb_model, unlist(test_x) %>% matrix(nrow = nrow(test_x))) %>% 
      {ifelse(. > 0.5, "Responder", "Non-responder")} %>%  # Cutoff at probability =  0.5
      as.factor()
    
    
    ## Confusion matrices
    glm_conmat[[fold]] = caret::confusionMatrix(data = glm_pred, 
                                                reference = test_y)
    
    rf_conmat[[fold]] = caret::confusionMatrix(data = rf_pred,
                                               reference = test_y)
    
    svm_conmat[[fold]] = caret::confusionMatrix(data = svm_pred,
                                                reference = test_y)
    
    xgb_conmat[[fold]] = caret::confusionMatrix(data = xgb_pred,
                                                reference = test_y)
    
    cat(paste("Completed: fold", fold))
  }
  
  
  ## Accuracies
  glm_acc[[repeats]] = sapply(glm_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  rf_acc[[repeats]] = sapply(rf_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  svm_acc[[repeats]] = sapply(svm_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  xgb_acc[[repeats]] = sapply(xgb_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  
  ## Balanced accuracies
  glm_balacc[[repeats]] = sapply(glm_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  rf_balacc[[repeats]] = sapply(rf_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  svm_balacc[[repeats]] = sapply(svm_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  xgb_balacc[[repeats]] = sapply(xgb_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  
  cat(paste("Completed: repeat", repeats))
}
```

```{r combination plotting training accuracies}

# Plotting



## Accuracy


## Getting data into appropriate form
comb_plot_data_acc = unlist(glm_acc) %>% 
  cbind(unlist(rf_acc),
        unlist(svm_acc),
        unlist(xgb_acc)) %>% 
  `colnames<-`(c("glm", "rf", "svm", "xgb")) %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 1:4, names_to = "method", values_to = "value")

## Boxplot accuracies
comb_training_acc_p = comb_plot_data_acc %>% 
  ggplot() +
  aes(x = method, y = value, colour = method) +
  labs(x = "", y = "Accuracy",
       title = "Combination Therapy: Model Accuracy on Training Data",
       subtitle = "5-fold repeated CV; n = 20, k = 5\n60 responders & 31 non-responders") +
  scale_x_discrete(labels = c("Logistic Regression", "RandomForest", "SVM", "XGBoost")) +
  geom_boxplot() +
  geom_jitter(width = 0.2, size = 0.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))



## Balanced Accuracy


## Getting data into appropriate form
comb_plot_data_balacc = unlist(glm_balacc) %>% 
  cbind(unlist(rf_balacc),
        unlist(svm_balacc),
        unlist(xgb_balacc)) %>% 
  `colnames<-`(c("glm", "rf", "svm", "xgb")) %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 1:4, names_to = "method", values_to = "value")

## Boxplot balanced accuracies
comb_training_balacc_p = comb_plot_data_balacc %>% 
  ggplot() +
  aes(x = method, y = value, colour = method) +
  labs(x = "", y = "Accuracy",
       title = "Combination Therapy: Model Balanced Accuracy on Training Data",
       subtitle = "5-fold repeated CV; n = 20, k = 5\n60 responders & 31 non-responders") +
  scale_x_discrete(labels = c("Logistic Regression", "RandomForest", "SVM", "XGBoost")) +
  geom_boxplot() +
  geom_jitter(width = 0.2, size = 0.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

comb_training_acc_p
comb_training_balacc_p

# ggsave("figures/combination_classification.jpg", plot = comb_training_balacc_p, height = 7, width = 10)
```


### Monotherapy

```{r monotherapy data balance}

table(barcodeprop_mono$response)
```

The data is completely balanced.


```{r monotherapy training accuracy}

## Initialising lists to store balanced accuracies
glm_acc = glm_balacc = rf_acc = rf_balacc = svm_acc = svm_balacc = xgb_acc = xgb_balacc = list()

set.seed(2021)

for(repeats in 1:20){
  
  ## Defining folds to use
  cvIndex = caret::createFolds(barcodeprop_mono$response, k = 10, returnTrain = TRUE)
  
  ## Initialising list of confusion matrices
  glm_conmat = rf_conmat = svm_conmat = xgb_conmat = list()
  
  for(fold in 1:5){
    
    ## Training data
    train_dat = barcodeprop_mono %>% .[cvIndex[[fold]], ] %>% select(-imageid, -batch)
    train_x = train_dat %>% select(-response)
    train_y = train_dat %>% pull(response)
    
    ## Testing data
    test_dat = barcodeprop_mono %>% .[-cvIndex[[fold]], ] %>% select(-imageid, -batch)
    test_x = test_dat %>% select(-response)
    test_y = test_dat %>% pull(response) 
    
    
    
    ## Fitting models
    glm_model = glm(response ~ ., 
                    data = train_dat, 
                    family = binomial(link = "logit"))
    
    rf_model = randomForest::randomForest(response ~ .,
                                          data = janitor::clean_names(train_dat)) # Need to clean names for randomforest to work
    
    svm_model = svm(train_x, train_y)
    
    xgb_model = xgboost(data = unlist(train_x) %>% matrix(nrow = nrow(train_x)),
                        label = as.numeric(train_y) - 1, # xgb needs label to be numeric 0 < y < 1
                        max_depth = 2, eta = 1, nthread = 1, nrounds = 2, objective = "binary:logistic")
    
    
    ## Predictions on test data
    glm_pred = predict(glm_model, test_x, type = "response") %>% 
      {ifelse(. > 0.5, "Responder", "Non-responder")} %>%  # Cutoff at probability =  0.5
      as.factor()
    
    rf_pred = predict(rf_model, janitor::clean_names(test_x), type = "response")
    
    svm_pred = predict(svm_model, test_x)
    
    xgb_pred = predict(xgb_model, unlist(test_x) %>% matrix(nrow = nrow(test_x))) %>% 
      {ifelse(. > 0.5, "Responder", "Non-responder")} %>%  # Cutoff at probability =  0.5
      as.factor()
    
    
    ## Confusion matrices
    glm_conmat[[fold]] = caret::confusionMatrix(data = glm_pred, 
                                                reference = test_y)
    
    rf_conmat[[fold]] = caret::confusionMatrix(data = rf_pred,
                                               reference = test_y)
    
    svm_conmat[[fold]] = caret::confusionMatrix(data = svm_pred,
                                                reference = test_y)
    
    xgb_conmat[[fold]] = caret::confusionMatrix(data = xgb_pred,
                                                reference = test_y)
    
    cat(paste("Completed: fold", fold))
  }
  
  
  ## Accuracies
  glm_acc[[repeats]] = sapply(glm_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  rf_acc[[repeats]] = sapply(rf_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  svm_acc[[repeats]] = sapply(svm_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  xgb_acc[[repeats]] = sapply(xgb_conmat, function(x) x$overall[["Accuracy"]]) %>% mean()
  
  ## Balanced accuracies
  glm_balacc[[repeats]] = sapply(glm_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  rf_balacc[[repeats]] = sapply(rf_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  svm_balacc[[repeats]] = sapply(svm_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  xgb_balacc[[repeats]] = sapply(xgb_conmat, function(x) x$byClass[["Balanced Accuracy"]]) %>% mean()
  
  cat(paste("Completed: repeat", repeats))
}
```

```{r monotherapy plotting training accuracies}

# Plotting



## Accuracy


## Getting data into appropriate form
mono_plot_data_acc = unlist(glm_acc) %>% 
  cbind(unlist(rf_acc),
        unlist(svm_acc),
        unlist(xgb_acc)) %>% 
  `colnames<-`(c("glm", "rf", "svm", "xgb")) %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 1:4, names_to = "method", values_to = "value")

## Boxplot accuracies
mono_training_acc_p = mono_plot_data_acc %>% 
  ggplot() +
  aes(x = method, y = value, colour = method) +
  labs(x = "", y = "Accuracy",
       title = "Monotherapy: Model Accuracy on Training Data",
       subtitle = "5-fold repeated CV; n = 20, k = 5\n53 responders & 53 non-responders") +
  scale_x_discrete(labels = c("Logistic Regression", "RandomForest", "SVM", "XGBoost")) +
  geom_boxplot() +
  geom_jitter(width = 0.2, size = 0.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))



## Balanced Accuracy --> not needed but included for completeness


## Getting data into appropriate form
mono_plot_data_balacc = unlist(glm_balacc) %>% 
  cbind(unlist(rf_balacc),
        unlist(svm_balacc),
        unlist(xgb_balacc)) %>% 
  `colnames<-`(c("glm", "rf", "svm", "xgb")) %>% 
  as.data.frame() %>% 
  pivot_longer(cols = 1:4, names_to = "method", values_to = "value")

## Boxplot balanced accuracies
mono_training_balacc_p = mono_plot_data_balacc %>% 
  ggplot() +
  aes(x = method, y = value, colour = method) +
  labs(x = "", y = "Accuracy",
       title = "Monotherapy: Model Balanced Accuracy on Training Data",
       subtitle = "5-fold repeated CV; n = 20, k = 5\n53 responders & 53 non-responders") +
  scale_x_discrete(labels = c("Logistic Regression", "RandomForest", "SVM", "XGBoost")) +
  geom_boxplot() +
  geom_jitter(width = 0.2, size = 0.2) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  theme_classic() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

mono_training_acc_p
mono_training_balacc_p

# ggsave("figures/monotherapy_classification.jpg", plot = mono_training_acc_p, height = 7, width = 10)
```


## Testing {.tabset}

Using a two-sample Hotelling's $t^2$ test with permutation (non-normally distributed data) and using Shaefer and Strimmer's James-Stein shrinkage estimator to calculate hte sample covariance matrices.

### Combination

```{r combination hotellings 2-sample permutation test}

# ## Setting seed for reproducibility
# set.seed(2021)
# 
# ## Obtaining 
# hotelling_comb = barcodeprop_comb %>% 
#   select(-imageid, -batch)
# 
# comb_proptest = hotelling.test(
#   .~response,
#   data = hotelling_comb,
#   shrinkage = TRUE,
#   perm = TRUE,
#   B = 10000
# )
# 
# save(comb_proptest, file = "rdata/comb_multivariate.Rdata")
load("rdata/comb_multivariate.Rdata")

comb_multivar_p = ggplot(data.frame(x = comb_proptest$results)) +
  aes(x = x) +
  geom_histogram(colour = "black", fill = "grey") +
  geom_vline(xintercept = comb_proptest$stats$statistic,
             colour = "red") +
  geom_text(mapping = aes(x = 40,
                          y = 100),
            label = paste("p =", comb_proptest$pval),
            colour = "red") +
  labs(x = expression(T^2),
       y = "Count",
       title = "Combination Therapy: Distribution of Permuted Test Statistics",
       subtitle = "Covariance matrices estimated w/ James-Stein estimator\nn = 1000 permutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# ggsave("figures/combination_multivariatetest.jpg", plot = comb_multivar_p, height = 7, width = 10)
```


There was no significant difference between cell proportions in responders and non-responders (p = `r proptest$pval`).

### Monotherapy

```{r monotherapy hotellings 2-sample permutation test}

# ## Setting seed for reproducibility
# set.seed(2021)
# 
# ## Obtaining 
# hotelling_mono = barcodeprop_mono %>% 
#   select(-imageid, -batch)
# 
# mono_proptest = hotelling.test(
#   .~response,
#   data = hotelling_mono,
#   shrinkage = TRUE,
#   perm = TRUE,
#   B = 10000
# )
# 
# save(mono_proptest, file = "rdata/mono_multivariate.Rdata")
load("rdata/mono_multivariate.Rdata")

mono_multivar_p = ggplot(data.frame(x = mono_proptest$results)) +
  aes(x = x) +
  geom_histogram(colour = "black", fill = "grey") +
  geom_vline(xintercept = mono_proptest$stats$statistic,
             colour = "red") +
  geom_text(mapping = aes(x = 40,
                          y = 100),
            label = paste("p =", mono_proptest$pval),
            colour = "red") +
  labs(x = expression(T^2),
       y = "Count",
       title = "Monotherapy: Distribution of Permuted Test Statistics",
       subtitle = "Covariance matrices estimated w/ James-Stein estimator\nn = 1000 permutations") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# ggsave("figures/monotherapy_multivariatetest.jpg", plot = mono_multivar_p, height = 7, width = 10)
```

There was a significant difference between cell proportions in responders and non-responders on anti-PD-1 monotherapy (p = `r proptest$pval`).

We can perform post-hoc univariate permutation tests to see which celltype in particular has an effect

```{r mono univariate tests}

## Setting seed for reproducibility
set.seed(2021)

## Defining number of permutations
B = 1000

## Initialising list to store values
mono_perm_tstat = list()

for(celltype in (barcodeprop_mono %>% select(-batch, -response, -imageid) %>% colnames())){
  
  permuted_dat = barcodeprop_mono
  t_null = vector("numeric", B)
  t_formula = as.formula(paste0("`", celltype, "` ", "~ response"))
  
  ## Obtaining true test statistic
  tt = t.test(t_formula, data = barcodeprop_mono)$statistic
  
  ## Obtaining permuted test statistics
  for(i in 1:B){
    permuted_dat$response = sample(barcodeprop_mono$response) #this is the permutation
    t_null[i] = t.test(t_formula, data = permuted_dat)$statistic
  }
  
  pval = mean(abs(t_null) >= abs(tt))
  
  mono_perm_tstat[[celltype]] = list(tt = tt, t_null = t_null, pval = pval)
  
}

mono_sig = mono_perm_tstat %>% sapply(function(x) x$pval) %>% sort() %>% 
  `*`(barcodeprop_mono %>% select(-batch, -response, -imageid) %>% colnames() %>% length()) %>%  # Bonferroni correction
  .[{which(. < 0.05)}]
mono_sig

# mono_univar1_hist_p = mono_perm_tstat[[names(mono_sig)[1]]][["t_null"]] %>% 
#   as.data.frame() %>% `colnames<-`(names(mono_sig)[1]) %>% 
#   ggplot() +
#   aes_string(x = paste0("`", names(mono_sig)[1], "`")) +
#   geom_histogram(colour = "black", fill = "grey", binwidth = 0.5) +
#   geom_vline(xintercept = mono_perm_tstat[[names(mono_sig)[1]]][["tt"]],
#              colour = "red") +
#   geom_text(mapping = aes(x = 8,
#                           y = 20),
#             label = paste("p =", mono_perm_tstat[[names(mono_sig)[1]]][["pval"]]),
#             colour = "red") +
#   labs(x = expression(T^2),
#        y = "Count",
#        title = "Monotherapy: Distribution of Permuted Test Statistics",
#        subtitle = paste("Comparing", names(mono_sig)[1], "proportions; Bonferroni adjusted pval")) +
#   theme_bw() +
#   theme(plot.title = element_text(hjust = 0.5),
#         plot.subtitle = element_text(hjust = 0.5))

mono_univar1_p = barcodeprop_mono %>% 
  ggplot() +
  aes(x = response, fill = response) +
  aes_string(y = paste0("`", names(mono_sig)[1], "`")) +
  geom_boxplot(alpha = 0.8) +
  geom_jitter(width = 0.2, cex = 0.5, alpha = 0.8) +
  labs(x = "", y = "Cell Proportion",
       title = paste("Proportion of", names(mono_sig)[1], "cells"),
       subtitle = "Monotherapy") +
  coord_flip() +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# ggsave("figures/monotherapy_univariatetest_1.jpg", plot = mono_univar1_p, width = 10, height = 7)
```







